---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/sebdalgarno/HGfetch.svg?branch=master)](https://travis-ci.org/sebdalgarno/HGfetch)

# haidawave

## Introduction

`haidawave` is as R package to calculate wave exposure (mean fetch) based on fetch and possibly wind data.
In addition to the `mean_fetch`, `wind_weights` and `weighted_fetch` functions, which are demonstrated below,
it also provides the `convert_proj` function to convert latitudes and longitudes
to eastings and northings in BC Albers (or a projection of the users choice);
the `nearest_point` function to join one data frame to another based on the shortest 
euclidean distance between their coordinates; and the `hg_fetch` and `hg_wind` 
to download the complete fetch and wind datasets for Haida Gwaii.

## Demonstration

```{r, message = FALSE}
library(haidawave)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(plyr)
library(dplyr)

# only consider points for Reef Island, Laskeek Bay, British Coluumbia
laskeek_fetch <- filter(haidawave::laskeek_fetch, Easting > 627000 & Northing < 890000)

# calculate mean fetch for each point
fetch <- ddply(laskeek_fetch, c("Easting", "Northing"), mean_fetch)

# Plot mean Fetch
gp <- ggplot(fetch) + geom_point(aes(x = Easting, y = Northing, color = Fetch), size = 0.5) +
  coord_fixed() +
  theme_few() +
  scale_color_distiller(palette = "Spectral")

print(gp)

# get wind weights by station and direction
cumshewa_wind <- ddply(haidawave::cumshewa_wind, c("Station"), wind_weights) %>% as.tbl()

cumshewa_wind

# averge fetch bearings by 10 degree increments so consistent with wind directions
laskeek_fetch %<>% mutate(Bearing = (Bearing + 5) %/% 10 * 10)

# bind wind weights
laskeek_fetch %<>% left_join(cumshewa_wind, by = c("Bearing" = "Direction"))

laskeek_fetch

# calculate mean weighted fetch
laskeek_fetch %<>% ddply(c("Easting", "Northing"), weighted_fetch) %>% as.tbl()

print(gp %+% laskeek_fetch)
```

## Installation

To install the latest version from GitHub
```
# install.packages("devtools")
devtools::install_github("sebdalgarno/haidawave")
```

## Contribution

Please report any [issues](https://github.com/sebdalgarno/haidawave/issues).

[Pull requests](https://github.com/sebdalgarno/haidawave/pulls) are always welcome.
